Methodology for Assessing the State of the Practice for Domain X

### 2021-01-09 StateOfPractice/Methodology/Methodology.tex
Ao suggested updates to methodology document

\begin{enumerate}
\item What artifacts are present in current software packages? 
\item What tools are used by current software packages?
\item What principles, processes, and methodologies are used in the development
  of current software packages?
\item What are the pain points for developers working on research software
  projects?  What aspects of the existing processes, methodologies and tools do
  they consider as potentially needing improvement?  How should processes,
  methodologies and tools be changed to improve software development and
  software quality?
\item For research software developers, what specific actions are taken to
  address the following:
\begin{enumerate}
\item usability
\item traceability
\item modifiability
\item maintainability
\item correctness
\item understandability
\item unambiguity
\item reproducibility
\item visibility/transparency
\end{enumerate} 
\item How does software designated as high quality by this methodology compare
  with top rated software by the community?
\end{enumerate}

@@

\begin{enumerate}
\item Identify the domain. (Section~\ref{SecIdentifyDomain})
\item \emph{Domain Experts}: Create a top ten list of software packages in the
  domain. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting
    Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Brief the Domain Experts on the overall objective, research proposal,
  research questions, measurement template, survey for short list projects,
  usability tests, maintainability
  experiments. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting
    Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Identify broad list of candidate software packages in the
  domain. (Section~\ref{SecIdentifyCandSoft})
\item Preliminary filter of software packages
  list. (Section~\ref{SecInitialFilter})
\item \emph{Domain Experts}: Review domain software
  list. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting
    Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Domain Analysis. (Section~\ref{SecDomainAnalysis})
\item \emph{Domain Experts}: Vet domain
  analysis. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting
    Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Gather source code and documentation for each prospective software
  package.
\item Collect empirical measures. (Section~\ref{SecEmpiricalMeasures})
\item Measure using measurement template. (Section~\ref{SecShallowMeasure})
\item Use AHP process to rank the software packages. (Section~\ref{SecAHP})
\item Identify a short list of top software packages, typically four to six, for
  deeper exploration according to the AHP rankings of the measurements.
\item \emph{Domain Experts}: Vet AHP ranking and short
  list. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting
    Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item With short list:
\begin{enumerate}
\item Survey developers
  (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Questions
    to Developers.pdf}{Questions to Developers})
\item Usability experiments
  (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}{Experiments})
\item Modifiability experiments
  (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}{Experiments})
\end{enumerate}
\item Rank short list. (Section~\ref{SecRankShortList}) \pmi{To be completed -
    add pairwise comparison tool}
\item Document answers for research questions.
\end{enumerate}

@@

\begin{enumerate}
\item The software functionality must fall within the identified domain.
\item The source code must be viewable.
\item The empirical measures listed in Section \ref{SecEmpiricalMeasures} should
  ideally be available, which implies a preference for GitHub repositories.
\item The software cannot be marked as incomplete or in an initial development
  phase.
\end{enumerate}

@@

Section \ref{rawdata} orients the reader on what raw data is collected.  Some of this data can be
observed from Git repository metrics. The rest can be collected using freeware
tools. \href{https://github.com/tomgi/git_stats}{GitStats} is used to measure
the number of binary files as well as the number of added and deleted lines in a
repository. This tool is also used to measure the number of commits over
different intervals of time. \href{https://github.com/boyter/scc}{Sloc Cloc and
  Code (scc)} is used to measure the number of text based files as well as the
number of total, code, comment, and blank lines in a repository. These tools
were selected due to their installability, usability, and ability to gather the
empirical measures listed below. Details on installing and running the tools can
be found in our 
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/A
  Guide to Empirical Measures.pdf} {Guide to Empirical Measures}. Section \ref{processeddata}
introduces the required processed data, which is calculated using the raw data.

@@

An AHP pairwise comparison between the short list packages with respect to
usability and modifiability survey results is conducted, and then reviewed by
the domain expert. The tool can be found here. The usability and modifiability
survey can be found
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}
{here}. This step is intended to start a conversation about the relative quality
of short-listed software packages.

>>>>

\begin{enumerate}
\item What artifacts are present in current software packages? 
\item What tools (development, dependencies, project management) are used by current software packages?
\item What principles, processes, and methodologies are used in the development
  of current software packages?
\item What are the pain points for developers working on research software
  projects?  What aspects of the existing processes, methodologies and tools do
  they consider as potentially needing improvement?  How should processes,
  methodologies and tools be changed to improve software development and
  software quality?
\item For research software developers, what specific actions are taken to
  address the following:
\begin{enumerate}
\item usability
\item traceability
\item modifiability
\item maintainability
\item correctness
\item understandability
\item unambiguity
\item reproducibility
\item visibility/transparency
\end{enumerate} 
\item How does software designated as high quality by this methodology compare
  with top rated software by the community?
\end{enumerate}

@@

\begin{enumerate}
\item Identify the domain. (Section~\ref{SecIdentifyDomain})
\item \emph{Domain Experts}: Create a top ten list of software packages in the
  domain. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting
    Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Brief the Domain Experts on the overall objective, research proposal,
  research questions, measurement template, survey for short list projects,
  usability tests, maintainability
  experiments. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting
    Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Identify broad list of candidate software packages in the
  domain. (Section~\ref{SecIdentifyCandSoft})
\item Preliminary filter of software packages
  list. (Section~\ref{SecInitialFilter})
\item \emph{Domain Experts}: Review domain software
  list. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting
    Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Domain Analysis. (Section~\ref{SecDomainAnalysis})
\item \emph{Domain Experts}: Vet domain
  analysis. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting
    Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item Gather source code and documentation for each prospective software
  package.
\item Collect empirical measures. (Section~\ref{SecEmpiricalMeasures})
\item Measure using measurement template. (Section~\ref{SecShallowMeasure})
\item Survey developers
(\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Questions
	to Developers.pdf}{Questions to Developers})
\item Use AHP process to rank the software packages. (Section~\ref{SecAHP})
\item Identify a short list of top software packages, typically four to six, for
  deeper exploration according to the AHP rankings of the measurements.
\item \emph{Domain Experts}: Vet AHP ranking and short
  list. (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Meeting
    Agenda withmDomainmExperts.pdf}{Meeting Agenda with Domain Experts})
\item With short list:
\begin{enumerate}
\item Usability experiments
  (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}{Experiments})
\item Modifiability experiments
  (\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}{Experiments})
\end{enumerate}
\item Rank short list. (Section~\ref{SecRankShortList}) \pmi{To be completed -
    add pairwise comparison tool}
\item Document answers for research questions.
\end{enumerate}

@@

\begin{enumerate}
\item The software functionality must fall within the identified domain.
\item The source code must be viewable.
\item The empirical measures listed in Section \ref{SecEmpiricalMeasures} should
  ideally be available, which implies a preference for GitHub-style repositories.
\item The software cannot be marked as incomplete or in an initial development
  phase.
\end{enumerate}

@@

Section \ref{rawdata} orients the reader on what raw data is collected.  Some of this data can be observed from GitHub repository metrics. The rest can be collected using freeware tools. \href{https://github.com/tomgi/git_stats}{GitStats} is used to measure
the number of binary files as well as the number of added and deleted lines in a
repository. This tool is also used to measure the number of commits over
different intervals of time. \href{https://github.com/boyter/scc}{Sloc Cloc and
  Code (scc)} is used to measure the number of text based files as well as the
number of total, code, comment, and blank lines in a repository. These tools
were selected due to their installability, usability, and ability to gather the
empirical measures listed below. Details on installing and running the tools can
be found in our 
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/A
  Guide to Empirical Measures.pdf} {Guide to Empirical Measures}. Section \ref{processeddata}
introduces the required processed data, which is calculated using the raw data.

@@

An AHP pairwise comparison between the short list packages with respect to
usability and modifiability experiment results is conducted, and then reviewed by the domain expert. The tool can be found here. The usability and modifiability
survey can be found
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/Methodology/Experiments.pdf}
{here}. This step is intended to start a conversation about the relative quality
of short-listed software packages.
